```{r}

```

---
title: "Cycle_trip"
author: "Kuro"
date: "2022-12-10"
output: html_document
---

Library install

```{r message=FALSE, warning=TRUE}
library(tidyverse)
library(ggplot2)
library(lubridate)
```

Import data

```{r}
q2_2019 <- read_csv("Divvy_Trips_2019_Q2.csv")
q3_2019 <- read_csv("Divvy_Trips_2019_Q3.csv")
q4_2019 <- read_csv("Divvy_Trips_2019_Q4.csv")
q1_2020 <- read_csv("Divvy_Trips_2020_Q1.csv")

```

WRANGLE DATA AND COMBINE INTO A SINGLE FILE

```{r}
colnames(q3_2019)
colnames(q4_2019)
colnames(q2_2019)
colnames(q1_2020)


```

```{r}
(q4_2019 <- rename(q4_2019
                   ,ride_id = trip_id
                   ,rideable_type = bikeid 
                   ,started_at = start_time  
                   ,ended_at = end_time  
                   ,start_station_name = from_station_name 
                   ,start_station_id = from_station_id 
                   ,end_station_name = to_station_name 
                   ,end_station_id = to_station_id 
                   ,member_casual = usertype))
(q3_2019 <- rename(q3_2019
                   ,ride_id = trip_id
                   ,rideable_type = bikeid 
                   ,started_at = start_time  
                   ,ended_at = end_time  
                   ,start_station_name = from_station_name 
                   ,start_station_id = from_station_id 
                   ,end_station_name = to_station_name 
                   ,end_station_id = to_station_id 
                   ,member_casual = usertype))

(q2_2019 <- rename(q2_2019
                   ,ride_id = "01 - Rental Details Rental ID"
                   ,rideable_type = "01 - Rental Details Bike ID" 
                   ,started_at = "01 - Rental Details Local Start Time"  
                   ,ended_at = "01 - Rental Details Local End Time"  
                   ,start_station_name = "03 - Rental Start Station Name" 
                   ,start_station_id = "03 - Rental Start Station ID"
                   ,end_station_name = "02 - Rental End Station Name" 
                   ,end_station_id = "02 - Rental End Station ID"
                   ,member_casual = "User Type"))

```

Seeing the changes

```{r}
str(q1_2020)
str(q4_2019)
str(q3_2019)
str(q2_2019)

```

Mutating the variable

```{r}
q4_2019 <-  mutate(q4_2019, ride_id = as.character(ride_id)
                   ,rideable_type = as.character(rideable_type)) 
q3_2019 <-  mutate(q3_2019, ride_id = as.character(ride_id)
                   ,rideable_type = as.character(rideable_type)) 
q2_2019 <-  mutate(q2_2019, ride_id = as.character(ride_id)
                   ,rideable_type = as.character(rideable_type)) 


```

Combining all the data

```{r}
all_trips <- bind_rows(q2_2019, q3_2019, q4_2019, q1_2020)
```

Removing some variable

```{r}
all_trips <- all_trips %>%  
  select(-c(start_lat, start_lng, end_lat, end_lng, birthyear, gender, "01 - Rental Details Duration In Seconds Uncapped", "05 - Member Details Member Birthday Year", "Member Gender", "tripduration"))

```

 CLEAN UP AND ADD DATA TO PREPARE FOR ANALYSIS

```{r}
colnames(all_trips)  #List of column names
nrow(all_trips)  #How many rows are in data frame?
dim(all_trips)  #Dimensions of the data frame?
head(all_trips)  #See the first 6 rows of data frame.  Also tail(all_trips)
str(all_trips)  #See list of columns and data types (numeric, character, etc)
summary(all_trips)  #Statistical summary of data. Mainly for numerics


```

\# There are a few problems we will need to fix:

\# (1) In the "member_casual" column, there are two names for members ("member" and "Subscriber") and two names for casual riders ("Customer" and "casual"). We will need to consolidate that from four to two labels.

\# (2) The data can only be aggregated at the ride-level, which is too granular. We will want to add some additional columns of data \-- such as day, month, year \-- that provide additional opportunities to aggregate the data.

\# (3) We will want to add a calculated field for length of ride since the 2020Q1 data did not have the "tripduration" column. We will add "ride_length" to the entire dataframe for consistency.

\# (4) There are some rides where tripduration shows up as negative, including several hundred rides where Divvy took bikes out of circulation for Quality Control reasons. We will want to delete these rides.

\
#1

```{r}
all_trips <-  all_trips %>% 
  mutate(member_casual = recode(member_casual
                           ,"Subscriber" = "member"
                           ,"Customer" = "casual"))

```

```{r}
table(all_trips$member_casual)
```

#2

```{r}
all_trips$date <- as.Date(all_trips$started_at) #The default format is yyyy-mm-dd
all_trips$month <- format(as.Date(all_trips$date), "%m")
all_trips$day <- format(as.Date(all_trips$date), "%d")
all_trips$year <- format(as.Date(all_trips$date), "%Y")
all_trips$day_of_week <- format(as.Date(all_trips$date), "%A")

```

Adding ride length

```{r}
all_trips$ride_length <- difftime(all_trips$ended_at,all_trips$started_at)
```

Converting the factor to numeric in ride_length

```{r}
is.factor(all_trips$ride_length)
all_trips$ride_length <- as.numeric(as.character(all_trips$ride_length))
is.numeric(all_trips$ride_length)

```

Removing bad data

```{r}
all_trips_v2 <- all_trips[!(all_trips$start_station_name == "HQ QR" | all_trips$ride_length<0),]
```

CONDUCT DESCRIPTIVE ANALYSIS

```{r}
mean(all_trips_v2$ride_length) #straight average (total ride length / rides)
median(all_trips_v2$ride_length) #midpoint number in the ascending array of ride lengths
max(all_trips_v2$ride_length) #longest ride
min(all_trips_v2$ride_length) #shortest ride

```

Comparing member and casual user

```{r}
aggregate(all_trips_v2$ride_length ~ all_trips_v2$member_casual, FUN = mean)
aggregate(all_trips_v2$ride_length ~ all_trips_v2$member_casual, FUN = median)
aggregate(all_trips_v2$ride_length ~ all_trips_v2$member_casual, FUN = max)
aggregate(all_trips_v2$ride_length ~ all_trips_v2$member_casual, FUN = min)

```

```{r}
aggregate(all_trips_v2$ride_length ~ all_trips_v2$member_casual + all_trips_v2$day_of_week, FUN = mean)
```

```{r}
all_trips_v2$day_of_week <- ordered(all_trips_v2$day_of_week, levels=c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday"))
```

```{r}
aggregate(all_trips_v2$ride_length ~ all_trips_v2$member_casual + all_trips_v2$day_of_week, FUN = mean)
```

```{r}
all_trips_v2 %>% 
  mutate(weekday = wday(started_at, label = TRUE)) %>%  #creates weekday field using wday()
  group_by(member_casual, weekday) %>%  #groups by usertype and weekday
  summarise(number_of_rides = n()							#calculates the number of rides and average duration 
  ,average_duration = mean(ride_length)) %>% 		# calculates the average duration
  arrange(member_casual, weekday)		
```

Visualizing

```{r}
all_trips_v2 %>% 
  mutate(weekday = wday(started_at, label = TRUE)) %>% 
  group_by(member_casual, weekday) %>% 
  summarise(number_of_rides =n()
            ,average_duration = mean(ride_length)) %>% 
  arrange(member_casual, weekday)  %>% 
  ggplot(aes(x = weekday, y = number_of_rides, fill = member_casual)) +
  geom_col(position = "dodge")
 
```

Visualization for average duration

```{r}
all_trips_v2 %>% 
  mutate(weekday = wday(started_at, label = TRUE)) %>% 
  group_by(member_casual, weekday) %>% 
  summarise(number_of_rides = n()
            ,average_duration = mean(ride_length)) %>% 
  arrange(member_casual, weekday)  %>% 
  ggplot(aes(x = weekday, y = average_duration, fill = member_casual)) +
  geom_col(position = "dodge")

```

```{r}
all_trips_v2 %>% 
  filter(member_casual == "member") %>% 
  group_by(start_station_name) %>% 
  summarise(number_of_rides = n(),
            average_duration = mean(ride_length)) %>%

  arrange(desc(number_of_rides))%>%

  head(30)
```

```{r}
all_trips_v2 %>% 
  filter(member_casual == "member") %>% 
  group_by(start_station_name, end_station_name) %>%
  summarise(number_of_rides = n(),
            average_duration = mean(ride_length)) %>%
  # Use the paste function to combine the start and end station name into a single column
  mutate(station = paste(start_station_name, end_station_name, sep = " to ")) %>%
  # Sort the results in descending order
  arrange(desc(number_of_rides)) %>%
  # Take only the top 10 results
  head(10) %>%
  # Create the plot
  ggplot(aes(x = number_of_rides, y = station)) +
  geom_bar(stat = "identity",position = "dodge")
```

```{r}
all_trips_v2 %>% 
  filter(member_casual == "casual") %>% 
  group_by(start_station_name, end_station_name) %>%
  summarise(number_of_rides = n(),
            average_duration = mean(ride_length)) %>%
  # Use the paste function to combine the start and end station name into a single column
  mutate(station = paste(start_station_name, end_station_name, sep = " to ")) %>%
  # Sort the results in descending order
  arrange(desc(number_of_rides)) %>%
  # Take only the top 10 results
  head(10) %>%
  # Create the plot
  ggplot(aes(x = number_of_rides, y = station)) +
  geom_bar(stat = "identity",position = "dodge")
```

EXPORT SUMMARY FILE FOR FURTHER ANALYSIS

```{r}
counts <- aggregate(all_trips_v2$ride_length ~ all_trips_v2$member_casual + all_trips_v2$day_of_week, FUN = mean)

```

```{r}

write.csv(counts, file = 'avg_ride_length.csv')

```
